---
title: "Hw8"
author: "Anish Mohan"
date: "March 8, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Summary
1. This file models production of Milk and Ice-Cream from 1995 to 2013.
2. The time series for milk and Ice cream production is not stationery.
3. There is significant seasonal component in the production of Ice cream and Milk
4. The best order of ARMA(p,q) is with p = 1 and q = 1
5. As the time moves into future confidence intervals are more wider.

## Modeling Milk Production

```{r}
#Reading the data
ca = read.csv("CADairyProduction.csv", stringsAsFactors = FALSE)
ca = na.omit(ca)
summary(ca)
str(ca)
milk.ts = ts(ca[,5], start = 1995, frequency = 12)

#ACF function
plot.acf <- function(df, col = 'remainder', is.df =TRUE){
  if(is.df) temp <- df[, col]
  else temp <- df
  par(mfrow = c(2,1))
  acf(temp, main = paste('ACF of', col))
  pacf(temp, main = paste('PACF of', col))
  par(mfrow = c(1,1))
}

#Decompositon function
ts.decomp <- function(df, col = 'milk.ts', span = 0.5, Mult = TRUE, is.df = TRUE){
  # if(Mult) temp = log(df[, col])  else temp = ts(df[, col]
  if(is.df) temp = log(df[, col])  
  else temp = df
  spans = span * length(temp)  
  fit <- stl(temp, s.window = "periodic", t.window = spans)
  plot(fit, main = paste('Decompositon of',col,'with lowess span = ', as.character(span)))
  fit$time.series
}

#Decomposition.
milk.decomp = ts.decomp(milk.ts, Mult = TRUE, is.df = FALSE, span = 0.5)
options(repr.pmales.extlot.width=8, repr.plot.height=6)
plot.acf(milk.decomp[, 3], is.df = FALSE)

#Arima model
ts.model = function(ts, col = 'remainder', order = c(0,0,1)){
  mod = arima(ts, order = order, include.mean = FALSE)
  print(mod)
  mod
}

#Trying different values of parameters.
milk.arima = ts.model(milk.decomp[, 3], col = 'ARIMA model for Milk Order', order = c(1,1,1))
milk.arima = ts.model(milk.decomp[, 3], col = 'ARIMA model for Milk Order', order = c(2,1,1))
milk.arima = ts.model(milk.decomp[, 3], col = 'ARIMA model for Milk Order', order = c(1,2,1))
milk.arima = ts.model(milk.decomp[, 3], col = 'ARIMA model for Milk Order', order = c(1,1,2))
milk.arima = ts.model(milk.decomp[, 3], col = 'ARIMA model for Milk Order', order = c(1,3,1))
milk.arima = ts.model(milk.decomp[, 3], col = 'ARIMA model for Milk Order', order = c(1,4,1))


#Forecasting the milk production
require(forecast)
fit.elect = auto.arima(milk.ts, max.p=3, max.q=3,
                       max.P=2, max.Q=2, max.order=5, max.d=2, max.D=1,
                       start.p=0, start.q=0, start.P=0, start.Q=0)
summary(fit.elect)

milk.forecast = forecast(fit.elect, h=12)
summary(milk.forecast)
plot(milk.forecast, xlab = "Year", ylab = "Milk Production")
```

## Modeling Ice-Cream Production

```{r}
#Reading the ice-cream production data.
ice.ts = ts(ca[,4], start = 1995, frequency = 12)
ice.decomp = ts.decomp(ice.ts, Mult = TRUE, is.df = FALSE, span = 0.5)
options(repr.pmales.extlot.width=8, repr.plot.height=6)
plot.acf(ice.decomp[, 3], is.df = FALSE)

#Arima model
ice.arima = ts.model(ice.decomp[, 3], col = 'ARIMA model for Ice-Cream Order', order = c(1,1,1))
ice.arima = ts.model(ice.decomp[, 3], col = 'ARIMA model for Ice-Cream Order', order = c(2,1,1))
ice.arima = ts.model(ice.decomp[, 3], col = 'ARIMA model for Ice-Cream Order', order = c(1,2,1))
ice.arima = ts.model(ice.decomp[, 3], col = 'ARIMA model for Ice-Cream Order', order = c(1,1,2))
ice.arima = ts.model(ice.decomp[, 3], col = 'ARIMA model for Ice-Cream Order', order = c(1,3,1))
ice.arima = ts.model(ice.decomp[, 3], col = 'ARIMA model for Ice-Cream Order', order = c(1,4,1))


require(forecast)
fit.elect = auto.arima(ice.ts, max.p=3, max.q=3,
                       max.P=2, max.Q=2, max.order=5, max.d=2, max.D=1,
                       start.p=0, start.q=0, start.P=0, start.Q=0)
summary(fit.elect)

#Forecasting for ice-cream production
ice.forecast = forecast(fit.elect, h=12)
summary(ice.forecast)
plot(ice.forecast, xlab = "Year", ylab = "Ice-Cream Production")
```


