---
title: "HW2"
author: "Anish Mohan"
date: "January 25, 2017"
output: html_document
---
##1. Summary and/or Key Conclusions

##2. Reviewing Data

  Source Data:-
  A. Tsanas, A. Xifara: 'Accurate quantitative estimation of energy performance of residential buildings using statistical machine learning tools', Energy and Buildings, Vol. 49, pp. 560-567, 2012


```{r}
library(ggplot2)
  # Read the dataset
  energy = read.csv("EnergyEfficiencyData.csv", header = TRUE, stringsAsFactors = FALSE)
```


```{r}
  # Review the names of the columns or the features
  names(energy)
```


```{r}
  # Print a summary of the datatypes in the data.
  str(energy)
```

  The data here shows that there are 768 rows (sample) and 10 columns (features).

```{r}
  # Quick check to see if there are any NA's in the data
  summary(is.na(energy))
```

  Per this chart, none of the columns have any NAs.
  
###. Exploratory Data Analysis

  1. Heating and Cooling Load as a function of Surface Area
```{r}
      #Scatter plot of the Heating/Cooling Load and Surface Area
      ggplot(energy, aes(y = Heating.Load, x = Surface.Area))+ geom_point() + xlab("Surface Area") + ylab("Heating Load") + ggtitle("Relationship between Heating Load and Surface Area")
      ggplot(energy, aes(y = Cooling.Load, x = Surface.Area))+ geom_point() + xlab("Surface Area") + ylab("Cooling Load") + ggtitle("Relationship between Cooling Load and Surface Area")
      
```

      + This scatter plot shows that that Heating/Cooling load is higher when the Surface areas is <675 units. As the Surface area increases beyond 675 units, the heating and cooling load starts decreasing
      
      + Next we will analyze if Surface Area in conjunction with other parameters impacts heating and cooling load


  2. Heating and Cooling Load as a function of Roof area along with impact of other variables
  
```{r}
      #Scatter Plot of Heating and Cooling Load and Surface Area with Orientation as a factor
      ggplot(energy, aes(y = Heating.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Roof.Area), alpha = 0.3)) + xlab("Surface Area") + ylab("Heating Load") + ggtitle("Impact of Surface Area and Roof.Area on Heating Load")
      
      ggplot(energy, aes(y = Cooling.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Roof.Area))) + xlab("Surface Area") + ylab("Cooling Load") + ggtitle("Impact of Surface Area and Roof.Area on Cooling Load")
```

    + In this plot it is shown that if Surface Area is larger than ~675 units, the Roof area is large. For such building, the Heating and the cooling load is lower than building that have small Surface area and Roof Area.
    
    
```{r}
      ggplot(energy, aes(y = Heating.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Relative.Compactness), alpha = 0.3)) + xlab("Surface Area") + ylab("Heating Load") + ggtitle("Impact of Surface Area and Relative Compactness on Heating Load ")
      
      ggplot(energy, aes(y = Cooling.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Relative.Compactness))) + xlab("Surface Area") + ylab("Cooling Load") + ggtitle("Impact of Surface Area and Relative Compactness on Cooling Load ")
```

  + In this plot is is shows that if Surface Area is larger than ~675 units, then the Relative Compactness is below 0.74 units and the heating and Cooling loads are lower.
    
    ```{r}
      ggplot(energy, aes(y = Heating.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Overall.Height), alpha = 0.3)) + xlab("Surface Area") + ylab("Heating Load") + ggtitle("Impact of Surface Area and Overall Height on Heating Load ")
      
      ggplot(energy, aes(y = Cooling.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Overall.Height))) + xlab("Surface Area") + ylab("Cooling Load") + ggtitle("Impact of Surface Area and Overall Height on Cooling Load ")
```

    + In this plot is is shows that if Surface Area is larger than ~675 units, then the Overall Height is ~3.5units and the heating and Cooling loads are lower.

    ```{r}
      ggplot(energy, aes(y = Heating.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Orientation))) + xlab("Surface Area") + ylab("Heating Load") + ggtitle("Impact of Surface Area and Orientation on Heating Load ")
      
      ggplot(energy, aes(y = Cooling.Load, x = Surface.Area)) + geom_point(aes(color = as.factor(Orientation))) + xlab("Surface Area") + ylab("Cooling Load") + ggtitle("Impact of Surface Area and Orientation on Cooling Load ")
```
    
    + This plot shows that there are no preferred orientation along with Surface Area such that it leads to lower heating or cooling load.  However, Direction = "2" is less common in lower Surface Area buildings where the Heating/Cooling load is larger 
